\vspace{-\baselineskip} Fast and scalable Gaussian process modeling with applications to astronomical time series \vspace{-2\baselineskip}

### 2017-06-04 paper/ms.tex
adding some words and the true values for the transit example

Simulation of a Gaussian process for large $N$ requires multiplying an
$N$-dimensional vector of Gaussian normal deviates, $\bf{q}$, by the square root
of the covariance matrix:
\begin{equation}
{\bf y} = K_\bvec{\alpha}^{1/2} {\bf q},
\end{equation}
where $q_i \sim N(0,1)$.  Note that $K_\bvec{\alpha} = K_\bvec{\alpha}^{1/2} \left(K_\bvec{\alpha}^{1/2}\right)^T$, which
comparing with the Cholesky factorization gives $K_\bvec{\alpha}^{1/2} = L D^{1/2}$,
where $(D^{1/2})_{ii} = \sqrt{D_{ii}}$.  Also note that the first term in $-2\ln{\mathcal{L}}$
with $\mu_\bvec{\theta} =0$ is given by
\begin{eqnarray}
\mathcal{Q} &=&{\bf y}^T K_\bvec{\alpha}^{-1} {\bf y}\\
&=& {\bf y}^T \left(K_\bvec{\alpha}^{-1/2}\right)^T K_\bvec{\alpha}^{-1/2} {\bf y}\\
&=& {\bf q}^T K_\bvec{\alpha}^{1/2} \left(K_\bvec{\alpha}^{-1/2}\right)^T K_\bvec{\alpha}^{-1/2} K_\bvec{\alpha}^{1/2} {\bf q}\\
&=& {\bf q}^T {\bf q}.
\end{eqnarray}
This last quantity is simply the dot product of an $N$-dimensional vector of independent standard normal random variables with itself,
which is distributed according to the chi-squared distribution with $N$ degrees of freedom, i.e.\ $\mathcal{Q} \sim \chi_N^2$.

@@

In this example, we inject the signal of a simulated exoplanet transit into a
real \kepler\ light curve and then demonstrate that we can recover the true
physical parameters of the exoplanet while modeling the stellar variability
using \celerite.
This example is different from all the previous examples~--~both simulated and
real~--~because, in this case, we are uninterested in the parameters of the
covariance model.
Instead, we're interested in inferring constraints on the parameters of the
mean model ($\mu_\bvec{\theta}(\bvec{x})$ from \sect{gps}), the model of the
signal caused by the transit of the planet.
We say ``mean model'' here to unambiguously distinguish it from the
``covariance model'', but this is what would generally be called the ``model''
in other analyses.
In this case, we have a light curve of a rotating star with a planet
transiting and we would like to fit for the physical parameters of the planet
while taking our uncertainty about the star's variability into account.
GP modeling has been used for this purpose throughout the exoplanet literature
\citep[for example][]{Dawson:2014, Barclay:2015, Evans:2015,
Foreman-Mackey:2016b, Grunblatt:2016}.

In \eq{gp-likelihood} the physical parameters of the planet are called
$\bvec{\theta}$ and, in this example, the mean function $\mu_\bvec{\theta}(t)$
is a limb-darkened transit light curve \citep{Mandel:2002} and the parameters
$\bvec{\theta}$ are a period $P$, a transit duration $T$, a phase or epoch
$t_0$, an impact parameter $B$, the radius of the planet in units of the
stellar radius $R_P/R_\star$, and several parameters describing the
limb-darkening profile of the star \citep{Claret:2011}.
As in the previous example, we model the stellar variability using a GP model
with a kernel given by \eq{rot-kernel} and fit for the parameters of the
planet $\bvec{\theta}$ and the stellar variability $\bvec{\alpha}$
simultaneously.

We take a month-long segment of the \kepler\ light curve for
KIC~1430163~--~the target from \sect{rotation}~--~and multiply it by a
simulated transit model with known parameters.
The fact that we are using a simulated transit enables us to confirm that we
correctly measure the planet parameters but we do not use this knowledge to
inform our fit.
The top panel of \Figure{transit-ml} shows the data including the simulated
transit.
The bottom panel shows the same data with the maximum likelihood prediction
for the stellar variation subtracted.
To find this model, we maximized the likelihood function in \eq{gp-likelihood}
with respect to the transit parameters $\bvec{\theta}$ and the variability
parameters $\bvec{\alpha}$ simultaneously.
\Figure{transit-corner} shows the marginalized posterior constraints on the
physical properties of the planet compared to the true values.
This procedure produces estimates of the planet parameters that are consistent
with the true values using \celerite\ as an effective model for the stellar
variability.

@@

\dfmtodo{Table 5 has two parameter with name "b":  the celerite coefficient,
and the impact parameter.}

\input{figures/transit/transit-params}

>>>>

Simulation of a Gaussian process for large $N$ requires multiplying an
$N$-dimensional vector of Gaussian normal deviates, $\bf{q}$, by the square root
of the covariance matrix:
\begin{equation}
{\bf y} = K_\bvec{\alpha}^{1/2} {\bf q},
\end{equation}
where $q_i \sim N(0,1)$.  Note that $K_\bvec{\alpha} = K_\bvec{\alpha}^{1/2} \left(K_\bvec{\alpha}^{1/2}\right)^T$, which
comparing with the Cholesky factorization gives $K_\bvec{\alpha}^{1/2} = L D^{1/2}$,
where $(D^{1/2})_{ii} = \sqrt{D_{ii}}$.  Also note that the exponent of
the likelihood with $\mu_\bvec{\theta} =0$ is given by
\begin{eqnarray}
Q &=&-\frac{1}{2}{\bf y}^T K_\bvec{\alpha}^{-1} {\bf y}\\
&=& -\frac{1}{2}{\bf y}^T \left(K_\bvec{\alpha}^{-1/2}\right)^T K_\bvec{\alpha}^{-1/2} {\bf y}\\
&=& -\frac{1}{2}{\bf q}^T K_\bvec{\alpha}^{1/2} \left(K_\bvec{\alpha}^{-1/2}\right)^T K_\bvec{\alpha}^{-1/2} K_\bvec{\alpha}^{1/2} {\bf q}\\
&=& -\frac{1}{2}{\bf q}^T {\bf q}.
\end{eqnarray}
The last quantity is simply the dot product of an $N$-dimensional vector of independent standard normal random variables with itself,
which is distributed according to the chi-squared distribution with $N$ degrees of freedom, i.e.\ $Q \sim \chi_N^2$.

@@

\response{
This example is different from all the previous examples~--~both simulated and
real~--~because, in this case, do not set the deterministic
mean function $\mu_\bvec{\theta}$ to zero.
Instead, we make inferences about $\mu_\bvec{\theta}$ because it is a
physically significant function and the parameters are fundamental properties
of the system.
This is an example using real data~--~because we use the light curve from
\sect{rotation}~--~but we multiply these data by a simulated transiting
exoplanet model with known parameters.
This allows us to show that we can recover the true parameters of the planet
even when the transit signal is superimposed on the real variability of a
star.
}
GP modeling has been used extensively for this purpose throughout the
exoplanet literature \citep[for example][]{Dawson:2014, Barclay:2015,
Evans:2015, Foreman-Mackey:2016b, Grunblatt:2016}.

In \eq{gp-likelihood} the physical parameters of the exoplanet are called
$\bvec{\theta}$ and, in this example, the mean function $\mu_\bvec{\theta}(t)$
is a limb-darkened transit light curve \citep{Mandel:2002} and the parameters
$\bvec{\theta}$ are the orbital period $P$, the transit duration $T$, the
phase or epoch $t_0$, the impact parameter $B$, the radius of the planet in
units of the stellar radius $R_P/R_\star$, and two parameters describing
the limb-darkening profile of the star \citep{Claret:2011,Kipping:2013}.
As in \sect{rotation}, we model the stellar variability using a GP model
with a kernel given by \eq{rot-kernel} and fit for the parameters of the
exoplanet $\bvec{\theta}$ and the stellar variability $\bvec{\alpha}$
simultaneously.

We take a month-long segment of the \kepler\ light curve for KIC~1430163 and
multiply it by a simulated transit model with known parameters.
The fact that we are using a simulated transit enables us to confirm that we
correctly measure the planet parameters but we do not use this knowledge to
inform our fit.
The top panel of \Figure{transit-ml} shows the data including the simulated
transit.
The bottom panel shows the same data with the maximum likelihood prediction
for the stellar variation subtracted.
To find this model, we maximized the likelihood function in \eq{gp-likelihood}
with respect to the transit parameters $\bvec{\theta}$ and the variability
parameters $\bvec{\alpha}$ simultaneously.
\Figure{transit-corner} shows the marginalized posterior constraints on the
physical properties of the planet compared to the true values.
This procedure produces estimates of the planet parameters that are consistent
with the true values using \celerite\ as an effective model for the stellar
variability.

@@

\input{figures/transit/transit-params-edited}
